---
title: "PML"
author: "Rohit Banerjee"
date: "April 15, 2018"
output: html_document
---

##Practical Machine Learning - Course Project

#Introduction

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
In this project, our goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. The goal of this project is to predict the manner in which they did the exercise.

#Loading the data set
```{r}
#Load libraries
library("caret")
trainingDataSet<- read.csv("pml-training.csv", sep=",", header=TRUE, na.strings = c("NA","",'#DIV/0!'))
testingDataSet<- read.csv("pml-testing.csv", sep=",", header=TRUE, na.strings = c("NA","",'#DIV/0!'))

```
#Cleaning the data
```{r}
trainingDataSet <- trainingDataSet[,(colSums(is.na(trainingDataSet)) == 0)]
testingDataSet <- testingDataSet[,(colSums(is.na(testingDataSet)) == 0)]
dim(testingDataSet)
dim(trainingDataSet)

```
Hence we have reduced down the columns in the data set.
#Preprocessing the data
```{r}
numericalsIdx <- which(lapply(trainingDataSet, class) %in% "numeric")

preprocessModel <-preProcess(trainingDataSet[,numericalsIdx],method=c('knnImpute', 'center', 'scale'))
pre_trainingDataSet <- predict(preprocessModel, trainingDataSet[,numericalsIdx])
pre_trainingDataSet$classe <- trainingDataSet$classe

pre_testingDataSet <-predict(preprocessModel,testingDataSet[,numericalsIdx])
nzv <- nearZeroVar(pre_trainingDataSet,saveMetrics=TRUE)
pre_trainingDataSet <- pre_trainingDataSet[,nzv$nzv==FALSE]

nzv <- nearZeroVar(pre_testingDataSet,saveMetrics=TRUE)
pre_testingDataSet <- pre_testingDataSet[,nzv$nzv==FALSE]
```

#Validation set

We want a 75% observation training dataset to train our model. We will then validate it on the last 70%.
```{r}
set.seed(12031987)
idxTrain<- createDataPartition(pre_trainingDataSet$classe, p=3/4, list=FALSE)
training<- pre_trainingDataSet[idxTrain, ]
validation <- pre_trainingDataSet[-idxTrain, ]
dim(training) ; dim(validation)


```
#Training the model

```{r}
library(randomForest)
modFitrf <- train(classe ~., method="rf", data=training, trControl=trainControl(method='cv'), number=5, allowParallel=TRUE, importance=TRUE )
modFitrf
```
#Interpretation
```{r}
varImpPlot(modFitrf$finalModel, sort = TRUE, type = 1, pch = 19, col = 1, cex = 0.6, main = "Importance of the Individual Principal Components")
```

#Cross Validation

```{r}
predValidRF <- predict(modFitrf, validation)
confus <- confusionMatrix(validation$classe, predValidRF)
confus$table
```



```{r}
accur <- postResample(validation$classe, predValidRF)
modAccuracy <- accur[[1]]
modAccuracy

out_of_sample_error <- 1 - modAccuracy
out_of_sample_error
```



The estimated accuracy of the model is 99.7% and the estimated out-of-sample error based on our fitted model applied to the cross validation dataset is 0.3%.



#Application of this model on the 20 test cases provided

```{r}
pred_final <- predict(modFitrf, pre_testingDataSet)
pred_final
```
These are our final results





